Chemistry of the Orphan Stream I: Identifying Stream Members \\ from Low-Resolution Spectroscopy

### 2012-09-19 2012-oss-mrs/Response_GDC_20120918.txt

Good morning Andy - here are what are hopefully my final comments on this paper.  Sure, it's taken a bit of too-ing and fro-ing but its a better paper for it.

Of course this version still doesn't have the couple of sentences that need to go in, probably in section 4.3, after the paragraph that ends "…fall within the shaded region illustrated in Figure 3." and before the para beginning "We have adopted…"

The text would be something like:
We have then investigated the status of these 10 stars in the PPMXL proper motion catalogue (ref), finding entries for 8 (?) of the candidates.  One star (OSS8) has a listed proper motion that is different from that of the others at the 3\sigma level [or whatever].  Consequently, we have reduced the membership likelihood  of this star from "High" to "Medium".

So then there should be 9 "High" classifications in Table 1, and the velocity measurements, etc, can note that they have been calculated for these 9 stars - removes any need discuss the fact that sigma goes down when OSS8 is excluded.  I think Table 1 should have a first column which gives an ID for the star - e.g. the OSS# we have used previously.  Saves quoting in the next paper complex ids like J1046219+004322!

OK, now specific comments.

1.  Author order.  We've never discussed this (well I don't recall any discussion) and normally I don't really care whether I am "et" or "al", but for this paper I do feel I have made quite a few contributions that have improved the science content.  So I wouldn't object to be listed second rather than third.  But you're in a better position to know whether that is appropriate or not, and I'm not going to lose any sleep over it if it remains as is.

2.  Abstract (yes, I know I'm being nit-picky again, sorry).
3rd line - you corrected Anglo, but not Astronomical as regards the telescope.  It's the Anglo-Australian Telescope.
3rd line - How about: "With modest S/N and" ==> "From modest S/N spectra and"
4th - 7th lines.  As it is now the "more metal-rich than previously found…" literally applies to the intrinsic dispersion.
I think it would be better to separate the mean and the dispersion sentences as, for example:
"Our data indicates… …stream metallicity of [Fe/H] = -1.71 +/- error, which is more metal-rich than previously found and unbiased by spectral type.  Further, the significant metallicity dispersion displayed by our probable members, 0.56 dex, suggests that the unidentified Orphan Stream parent was a dSph satellite."
(note that the error in the mean metallicity could simply be 0.56/sqrt(9) = 0.19, i.e., standard error of the mean.  But we definitely need to give some number for the uncertainty in the mean metallicity).
9th-10th lines.  The 0.2 +/- 3.1 kms-1 number is the estimate of the intrinsic velocity dispersion, not the observed dispersion.  I think you could drop the number and just say "The observed velocity dispersion….  is consistent with arising from the velocity uncertainties alone."
11th line "This is indicative" ==> "This indicates".

3.  Discussion at the end of 4.3.  You have "…it is clear that the four most probable members lie almost precisely on this isochrone".  I think this is a left-over from the previous approach.  We now have 9 most probable members and only a subset are precisely on this isochrone - the more metal-poorer and more metal-richer aren't.  So this bit needs slight revision - something like with this mean metallicity and adopted distance, 4 of the 9 stars are on the isochrone.  Or may drop it altogether?  Similarly the bit in Section 5 "our most probable members lay perfectly on a 10 Gys isochrone of [Fe/H] = -1.71" needs either revision or dropping.

4. Other minor things.
5 lines up from para starting "In this letter", right hand column page 1.  "full spatial extent of the stream" ==> "full spatial extent of the Orphan Stream"
2nd line, left column page 3.  "criterion outlined" ==> "criterion as outlined"
7th line, same place.  "There is no sharp kinematic" ==> "There is no obvious sharp kinematic"
next line - ditch "exists" - not required.
2nd last line before 4.2.  Change "kinematics" to "velocities"
1st line of 4.2 "widths" ==> "width"  (the line has only one width).
Second paragraph of Section 5.  You have two sentences both starting "If the…".  I suggest that maybe we could just drop the second one - i.e. simply remove the sentence "If the stream density… …Stripe 1540 field".  Reason - well it doesn't actually contain much information, and is largely implicit in what's said in the previous sentence.
Third para of Section 5. - again there's confusion between the intrinsic dispersion (i.e. with the contribution from the errors removed) and the observed dispersion, which contains the contribution from the velocity errors.  So I would suggest you use "intrinsic dispersion" for the 0.2 +/- 3.1 kms-1 [or whatever figure finally goes here] and drop the "i.e. comparable to our velocity resolution".  Or say that the observed dispersion is dominated by the velocity errors, indicating that the intrinsic dispersion is small.

Finally, if it doesn't kill us for space, can we add in the acknowledgements:
"GDaC is also grateful for the support received during an extended visit to the Institute of Astronomy, University of Cambridge, during which this work was completed."
(I'd like to acknowledge the support I have received here and this paper has the best combination of work I've done here and likelihood of being published on a reasonable timescale).

Cheers,
Gary

### 2012-09-19 2012-oss-mrs/draft-oss-mrs.tex
Incorporated final GDC comments

%!TEX TS-program = pdflatex
\documentclass{emulateapj}

\usepackage{amssymb} % for resolution symbol

@@

\author{Andrew R. Casey\altaffilmark{1,2}, Stefan C. Keller\altaffilmark{1}, Gary Da Costa\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}
\altaffiltext{2}{Massachusetts Institute of Technology, Kavli Institute for Astrophysics and Space Research,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\begin{abstract}
We present coordinates and properties of candidate K giant members in the Orphan Stream, which have been identified from low-resolution spectroscopic data taken with AAOmega on the Anglo-Astronomical Telescope. With modest S/N and independent cuts in photometry, kinematics, gravity and metallicity we yield self-consistent, highly probable stream members. Our data indicates an overall stream metallicity of [Fe/H] $= -1.71$ with an intrinsic dispersion of $\sigma = 0.56$\,dex which is more metal-rich than previously found and unbiased by spectral type. The metallicity dispersion in our stream sample suggests that the unidentified Orphan Stream parent is a dSph satellite. We find a revised stream distance of $23.5 \pm 1.0$\,kpc near the celestial equator, and our kinematic signature peaks at $V_{GSR} = 00.0$\,km s$^{-1}$. The observed velocity dispersion of our highest probability members is consistent with arising from the velocity uncertainties alone, indicating that at least along the line-of-sight, the Orphan Stream is kinematically quite cold. We highlight likely members for high-resolution spectroscopic follow-up.
\end{abstract}

@@

The Milky Way stellar halo has partly formed through the accretion of satellites which are disrupted by tidal forces as they fall into the Milky Way's potential. Stars which were once gravitationally bound to the satellite are left behind in a stream of stars. Stream kinematics are sensitive to the shape of the galactic potential, allowing us to constrain the Milky Way potential, and reconstruct the formation history of the Milky Way. The level of accreted substructure in the Milky Way has only recently become apparent through multi-band photometric surveys like the Sloan Digital Sky Survey (SDSS). The more prominent of the detectable substructures, like Sagittarius, have been well-studied. One of the more prominent \--- yet less studied \--- substructures is that of the Orphan Stream.

@@

As \citet{Sales;et-al_2008} point out, there is a natural observational bias towards more massive and recent mergers like Sagittarius. Consequently, the fainter end of this substructure distribution has yet to be fully recovered, or thoroughly examined. Interestingly, there are indications that  some fainter substructures like the Orphan Stream and the Palomar 5 tidal tails \citep{Odenkirchen;et-al_2009} have orbits which seem to be best-fit by Milky Way models with nearly 60\% less mass \citep{Newberg;et-al_2010} than generally reported by \citet{Xue;et-al_2008} and \citet{Koposov;et-al_2010}. Such a discrepancy in the mass of the Milky Way is troublesome. More complete photometric and kinematic maps of these low total luminosity streams may provide the best test as to whether this mass discrepancy is real, or an artefact of incomplete observations. Whilst the full spatial extent of the stream remains incomplete, we can examine the detailed chemistry of its members and investigate the stream history, as well as make predictions about the nature of the host system.

In this letter we present a detailed, self-consistent analysis to identify K giant members of the Orphan Stream. Using our selection method we have catalogued the locations of ten highly probable Orphan Stream candidates, all worthy of high-resolution spectroscopic follow up. In the following section we outline our photometric target selection. In \S\ref{sec:observations} we describe the low-resolution spectroscopic observations. The data analysis, including stream identification, is discussed in \S\ref{sec:analysis} and in \S\ref{sec:conclusions} the conclusions, predictions and future work are presented.

@@

The Orphan Stream is closest to us in two locations on the edge of the SDSS footprint: at the celestial equator \citep{Belokurov;et-al_2007}, and along outrigger SEGUE Stripe 1540 \citep{Newberg;et-al_2010}. These two locations are unequivocally the best place to recover bright stream members. We have targeted two fields centered on ($\alpha$, $\delta$) = (10:48:15, 00:00:00) and (10:48:15, $-$02:30:00), and employed a combination of colour cuts with the SDSS DR 7 \citep{Abazajian;et-al_2009} data set in order to identify likely K giants:
\begin{eqnarray}
0.6 <& (g-i)_0 &< 1.7 \\
-15(g-i)_0 + 27 <& g_0 &< -3.75(g-i)_0 + 22.5 \\
15  <& i_0  &< 18 
\end{eqnarray}

@@

Observations took place on the Anglo-Astronomical Telescope using the AAOmega spectrograph in April 2009. AAOmega is a fibre-fed, dual beam multi-object spectrograph which is capable of simultaneously observing spectra of 392 (science and sky) targets across a $2^\circ$ field of view. We used the 5700\,{\AA} dichroic in combination with the 1000I grating in the red arm, and the 580V grating in the blue arm. This provides a spectral coverage between $800 \leq \lambda \leq 950$\,nm in the red at $\mathcal{R} \approx 4400$, and between $370 \leq \lambda \leq 580$\,nm with a lower resolving power of $\mathcal{R} \approx 1300$ in the blue.

@@

Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin;et-al_2003}. We have selected predicted stars from the Besan\c{c}on model using the same criterion outlined in \S\ref{sec:target-selection} after using the \citet{Jordi;et-al_2006} colour transformations. It is clear that our target selection has yielded mostly nearby thick disk dwarf stars with $V_{GSR} \approx -120$\,km s$^{-1}$. The expected kinematic signature for the Orphan Stream, as \citet{Newberg;et-al_2010} reported, is labelled in Figure \ref{fig:velocities}. There is no sharp kinematic peak representative of the Orphan Stream exists in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution.

In this region \citet{Newberg;et-al_2010} found the Orphan Stream to have $V_{GSR} = 110$ km s$^{-1}$ from BHB stars, approximately $105$\,km s$^{-1}$ on our scale. To isolate potential Orphan Stream members we have nominated a relatively wide kinematic selection region between $75 \leq V_{GSR} \leq 135$ km s$^{-1}$, which yields twenty-five Orphan Stream candidates. The typical uncertainty in our kinematics is $\pm{}5.0$\,km s$^{-1}$, although for our fainter targets this quickly increases above 10\,km s$^{-1}$.

@@

We have measured the equivalent widths of the gravity-sensitive Mg\,\textsc{I} line at 8807 \AA{} to distinguish dwarfs from giants \citep{Battaglia;Starkenburg_2012}. At a given temperature (or $g - r$) and metallicity, giant stars present narrower Mg\,\textsc{I} absorption lines than their dwarf contaminants. Given our target selection, our sample is likely to contain many more dwarfs than giants. In some cases no Mg\,\textsc{I} 8807 \AA{} line was apparent, so an upper limit was estimated based on the S/N. In these cases the candidate was considered a ``non-dwarf" because we cannot exclusively rule out a metal-poor sub-giant with this criterion alone. For these purposes we are only looking for a simple indication as to whether a star is likely a dwarf or not. Figure \ref{fig:ew-mg} illustrates the trend with EW$_{\lambda8807}$ against SDSS de-reddened\footnote{All magnitudes presented in this letter are de-reddened using the \citet{Schlegel;Finkbeiner;Davis_1998} dust maps.} $g - r$, illustrating the dominant upper dwarf branch we wish to exclude. The adopted dwarf/giant separation line is shown. Admittedly this selection criterion is not particularly strict, but we are allowing multiple inclusive cuts and taking the intersection of these criteria to identify Orphan Stream giants.

@@

We have measured the equivalent widths of the Ca\,\textsc{II} near-infrared triplet lines for stars that meet our kinematic and surface gravity criteria. After correcting for luminosity, the equivalent width of the Ca\,\textsc{II} triplet lines provide a good indication of the overall metallicity of a RGB star \citep{Amandroff;Da_Costa_1991}. We have employed the \citet{Starkenburg;et-al_2010} relationship and corrected for luminosity in $g$ against the horizontal branch magnitude at $g_{HB}$ = 17.1 \citep{Newberg;et-al_2010}. Strictly speaking, the Ca\,\textsc{II}\---[Fe/H] calibration is only valid for stars brighter than the horizontal branch, although the relationship only becomes significantly inappropriate near $g \-- g_{HB} \sim +1$ \citep{Saviane;et-al_2012}. Many of our candidates are fainter than this valid luminosity range, and therefore they should not be excluded solely because of their derived metallicities. Stars fainted than $g_{HB}$ will have slightly lower metallicities than predicted by our Ca\,\textsc{II}\---[Fe/H] relationship, and for these stars we will only use metallicities to assign a relative qualitative likelihood for stream membership.
Given a distance to the Orphan Stream, we can also estimate a star's metallicity through isochrone fitting. We have used a 10\,Gyr \citet{Girardi;et-al_2008} isochrone at 21.4\,kpc \citep{Newberg;et-al_2010} and found metallicities for all eighteen likely stream members from their best-fitting isochrone (see Figure \ref{fig:feh}). Derived metallicities from Ca\,\textsc{II} line strengths and isochrone fitting that are consistent (within $\pm0.3$\,dex) indicates these measurements are reliable, and that these stars are indeed at a distance of $\sim$21.4\,kpc. We find ten highly likely stream members with consistently derived metallicities. They fall within the shaded region illustrated in Figure \ref{fig:feh}.

We have adopted a final overall metallicity for each star based on the quality of our [Fe/H] measurements. Adopted values are tabulated in Table \ref{tab:oss-members}. From our ten highly likely stream members we find an overall stream metallicity of $[\mbox{Fe/H}] = -1.71$ with a dispersion of $\sigma = 0.56$ dex. This abundance spread is larger than typically seen in globular cluster stars and is more representative of the chemical spreads seen in dSph satellites. \citet{Newberg;et-al_2010} found a stream metallicity of $[\mbox{Fe/H}] = -2.1$ from BHB stars, which is not inconsistent with our measurement given the spread in Orphan Stream K-giant metallicities.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/feh.pdf}
	\caption{Metallicities from the Ca\,\textsc{II} triplet lines versus those found from fitting isochrones to the eighteen stars that meet our kinematic and surface gravity criteria. Both abundance determinations imply these stars are RGB members of the Orphan Stream at a distance of $\sim$21.4\,kpc \citep{Newberg;et-al_2010}. Consistency between these methods indicates highly likely stream membership (green region). The minimum isochrone metallicity, and a representative uncertainty of 0.2\,dex for abundance measurements is shown.}
	\label{fig:feh}
\end{figure}

Given that our Orphan Stream giants cover a wide evolutionary range along the giant branch (Figure \ref{fig:cmd}), we are in a good position to revise the distance estimate to the stream. Given a 10\,Gyr \citet{Girardi;et-al_2008} isochrone at $[\mbox{Fe/H}] = -1.71$, we find a best-fitting distance to the stream of $23.5 \pm 1.0$\,kpc at ($l, $b) = ($250\,^\circ$, $50\,^\circ$). This isochrone is also shown in Figure \ref{fig:cmd}, and it is clear that the four most probable member stars lie almost precisely on this isochrone. Our derived distance is slightly farther than the $21.4 \pm 1.0$\,kpc found by \citet{Newberg;et-al_2010}, although given the difficulty in measuring distances these values are largely in agreement.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/cmd.pdf}
	\caption{Color magnitude diagram showing our observed candidates (grey). Observations fulfilling kinematic and gravity cuts are colored by their metallicity, and those with upper limits for surface gravity are marked as triangles ($\triangledown$). Highly probable stream members with consistent line strength- and isochrone-derived metallicities are circled. Relevant 10\,Gyr \citet{Girardi;et-al_2008} isochrones at $[\mbox{Fe/H}] = -1.5$ (dotted), $-2.0$ (dashed) at 21.4\,kpc \citep{Newberg;et-al_2010} are shown, as well as our best-fitting 10\,Gyr isochrone of $[\mbox{Fe/H}] = -1.71$ at 23.5\,kpc (solid).}
	\label{fig:cmd}
\end{figure}

\begin{deluxetable*}{ccccccccccccc}
\tabletypesize{\scriptsize}
\tablecaption{Identified Orphan Stream Candidates\label{tab:oss-members}}
\tablehead{
	\colhead{$\alpha$} &
	\colhead{$\delta$} &
	\colhead{$g$} &
	\colhead{$g \-- r$} &
	\colhead{$V_{hel}$} &
	\colhead{$V_{GSR}$} &
	\colhead{$V_{err}$} &
	\colhead{$EW_{\lambda8807}$} &
	\colhead{[Fe/H]$_{Ca}$} &
	\colhead{[Fe/H]$_{iso}$} &
	\colhead{[Fe/H]\tablenotemark{a}} &
	\colhead{Stream} \\
(J2000) & (J2000) & & & (kms $^{-1}$) & (km s$^{-1}$) & (km s$^{-1}$) & (\AA) & (dex) & (dex) & (dex) & Prob.
}
\startdata
10:46:21.9 & $+$00:43:21.8 & 17.52 & 0.50 & 209.2 &  73.3 &  9.3 & 0.273 & --1.78 &$<$--2.28& --1.78 & Low \\
10:46:29.3 & $-$00:19:38.5 & 17.77 & 0.56 & 217.4 &  78.4 &  5.2 & 0.126 & --1.63 & --1.68  & --1.63 & High \\
10:46:50.4 & $-$00:13:15.6 & 17.33 & 0.51 & 215.5 &  77.0 &  4.0 & 0.416 & --1.31 &$<$--2.28& --1.31 & Low \\
10:47:06.1 & $-$01:56:03.9 & 18.74 & 0.54 & 218.4 &  74.9 & 17.6 & 0.452 &:--1.12\tablenotemark{b}& --1.40  & --1.40 & High \\
10:47:15.0 & $-$03:15:03.9 & 18.66 & 0.54 & 256.8 & 109.5 &  9.0 &$<$0.19&:--1.85\tablenotemark{b}& --1.43  & --1.43 & Medium \\
10:47:17.6 & $+$00:25:07.7 & 16.09 & 0.72 & 215.7 &  79.2 &  3.3 & 0.212 & --1.84 & --1.80  & --1.84 & High \\
10:47:29.1 & $-$02:02:22.6 & 17.86 & 0.47 & 236.9 &  93.2 & 29.8 &$<$0.40& --2.82 &$<$--2.28& --2.82 & High \\
10:47:30.1 & $-$00:01:24.5 & 17.25 & 0.61 & 221.3 &  83.6 &  3.5 & 0.123 & --1.62 & --1.68  & --1.62 & High \\
10:48:20.9 & $+$00:26:34.4 & 17.88 & 0.55 & 254.9 & 118.9 & 11.7 & 0.467 & --1.65 & --1.73  & --1.65 & High \\
10:48:27.8 & $+$00:55:24.0 & 17.72 & 0.51 & 259.1 & 124.5 &  6.7 & 0.182 & --1.48 &$<$--2.28& --1.48 & Low \\
10:48:31.9 & $+$00:03:35.7 & 17.02 & 0.58 & 242.3 & 105.1 &  5.1 & 0.234 & --1.12 & --2.10  & --1.12 & Low \\
10:48:44.4 & $-$02:53:08.8 & 18.35 & 0.62 & 253.9 & 108.2 &  9.0 & 0.183 &:--1.01\tablenotemark{b}& --1.17  & --1.17 & High \\
10:48:46.9 & $-$00:32:27.8 & 17.85 & 0.46 & 248.2 & 109.3 &  8.1 & 0.324 & --2.37 &$<$--2.28& --2.37 & High \\
10:49:08.3 & $+$00:02:00.2 & 16.27 & 0.62 & 218.5 &  81.5 &  4.6 & 0.034 & --2.70 &$<$--2.28& --2.70 & High \\
10:50:13.1 & $+$00:33:52.7 & 16.13 & 0.58 & 229.7 &  94.7 &  5.1 & 0.391 & --1.54 &$<$--2.28& --1.54 & Low \\
10:50:24.2 & $-$01:49:05.4 & 17.94 & 0.54 & 252.0 & 109.9 & 25.4 & 0.151 & --1.06 & --1.73  & --1.06 & Low \\
10:50:33.8 & $+$00:12:19.1 & 17.82 & 0.60 & 233.5 &  97.5 &  5.9 & 0.596 & --0.90 & --1.43  & --0.90 & Medium \\
10:51:19.7 & $+$00:05:15.5 & 17.81 & 0.65 & 218.7 &  82.7 &  5.0 & 0.198 & --1.16 & --1.20  & --1.16 & High 
\enddata
\tablenotetext{a}{Final adopted [Fe/H] value based on quality of two metallicity measurements.}
\tablenotetext{b}{Sufficiently fainter than $g_{HB}$ to qualify this measurement as uncertain.}
\end{deluxetable*}

@@

\bibliographystyle{apj}
\bibliography{bibliography}
\end{document}

>>>>

%!TEX TS-program = pdflatex
\documentclass{emulateapj}
\usepackage{amssymb} % for resolution symbol

@@

\author{Andrew R. Casey\altaffilmark{1,2}, Gary Da Costa\altaffilmark{1}, Stefan C. Keller\altaffilmark{1}, Elizabeth Maunder\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; \email{acasey@mso.anu.edu.au}}
\altaffiltext{2}{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, 77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\begin{abstract}
We present coordinates and properties of candidate K giant members in the Orphan Stream, which have been identified from low-resolution spectroscopic data taken with AAOmega on the Anglo-Australian Telescope. From modest S/N spectra and independent cuts in photometry, kinematics, gravity and metallicity we yield self-consistent, highly probable stream members. We find a revised stream distance of $22.5\,\pm\,1.0$\,kpc near the celestial equator, and our kinematic signature peaks at $V_{GSR} = 82.1\,\pm\,1.4$\,km s$^{-1}$. The observed velocity dispersion of our most probable members is consistent with arising from the velocity uncertainties alone. This indicates that at least along the line-of-sight, the Orphan Stream is kinematically quite cold. Our data indicates an overall stream metallicity of [Fe/H] $= -1.63\,\pm\,0.19$\,dex which is more metal-rich than previously found and unbiased by spectral type. Furthermore, the significant metallicity dispersion displayed by our probable members, $\sigma = 0.56$\,dex, suggests that the unidentified Orphan Stream parent was a dSph satellite. We highlight likely members for high-resolution spectroscopic follow-up.
\end{abstract}

@@

The Milky Way stellar halo has partly formed through the accretion of satellites which are disrupted by tidal forces as they fall into the Milky Way's potential. Stars which were once gravitationally bound to the satellite are left behind in a stream of stars. Stream kinematics are sensitive to the shape of the galactic potential, allowing us to constrain the Milky Way potential, and reconstruct the formation history of the Galaxy. The level of accreted substructure in the Milky Way has only recently become apparent through multi-band photometric surveys like the Sloan Digital Sky Survey (SDSS). The more prominent of the detectable substructures, like Sagittarius, have been well-studied. One of the more prominent \--- yet less studied \--- substructures is that of the Orphan Stream.

@@

As \citet{Sales;et-al_2008} point out, there is a natural observational bias towards more massive and recent mergers like Sagittarius. Consequently, the fainter end of this substructure distribution has yet to be fully recovered, or thoroughly examined. Interestingly, there are indications that  some fainter substructures like the Orphan Stream and the Palomar 5 tidal tails \citep{Odenkirchen;et-al_2009} have orbits which seem to be best-fit by Milky Way models with nearly 60\% less mass \citep{Newberg;et-al_2010} than generally reported by \citet{Xue;et-al_2008} and \citet{Koposov;et-al_2010}. Such a discrepancy in the mass of the Milky Way is troublesome. More complete photometric and kinematic maps of these low total luminosity streams may provide the best test as to whether this mass discrepancy is real, or an artefact of incomplete observations. Whilst the full spatial extent of the Orphan Stream remains incomplete, we can examine the detailed chemistry of its members and investigate the stream history, as well as make predictions about the nature of the host system.

In this letter we present a detailed, self-consistent analysis to identify K giant members of the Orphan Stream. Using our selection method we have catalogued the locations of nine highly probable Orphan Stream candidates, all worthy of high-resolution spectroscopic follow up. In the following section we outline our photometric target selection. In \S\ref{sec:observations} we describe the low-resolution spectroscopic observations. The data analysis, including stream identification, is discussed in \S\ref{sec:analysis} and in \S\ref{sec:conclusions} the conclusions, predictions and future work are presented.

@@

The Orphan Stream is closest to us in two locations on the edge of the SDSS footprint: at the celestial equator \citep{Belokurov;et-al_2007}, and along outrigger SEGUE Stripe 1540 \citep{Newberg;et-al_2010}. These two locations are unequivocally the best place to recover bright stream members. We have targeted two fields centered on $(\alpha, \delta) =$ (10:48:15, 00:00:00) and (10:48:15, $-$02:30:00), and employed a combination of colour cuts with the SDSS DR 7 \citep{Abazajian;et-al_2009} data set in order to identify likely K giants:
\begin{eqnarray}
0.6 <& (g-i)_0 &< 1.7 \\
-15(g-i)_0 + 27 <& g_0 &< -3.75(g-i)_0 + 22.5 \\
15  <& i_0  &< 18 
\end{eqnarray}

@@

Observations took place on the Anglo-Australian Telescope using the AAOmega spectrograph in April 2009. AAOmega is a fibre-fed, dual beam multi-object spectrograph which is capable of simultaneously observing spectra of 392 (science and sky) targets across a $2\,^\circ$ field of view. We used the 5700\,{\AA} dichroic in combination with the 1000I grating in the red arm, and the 580V grating in the blue arm. This provides a spectral coverage between $800 \leq \lambda \leq 950$\,nm in the red at $\mathcal{R} \approx 4400$, and between $370 \leq \lambda \leq 580$\,nm with a lower resolving power of $\mathcal{R} \approx 1300$ in the blue.

@@

Figure \ref{fig:velocities} shows a histogram of our galactocentric velocities, compared to the predicted smooth line-of-sight velocity distribution for this region from the Besan\c{c}on model \citep{Robin;et-al_2003}. We have selected predicted stars from the Besan\c{c}on model using the same criterion as outlined in \S\ref{sec:target-selection} after using the \citet{Jordi;et-al_2006} colour transformations. It is clear that our target selection has yielded mostly nearby thick disk dwarf stars with $V_{GSR} \approx -120$\,km s$^{-1}$. The expected kinematic signature for the Orphan Stream, as \citet{Newberg;et-al_2010} reported, is labelled in Figure \ref{fig:velocities}. There is no obvious sharp kinematic peak representative of the Orphan Stream in our sample. From kinematics alone, our targets appears largely indistinguishable from a smooth halo distribution.

In this region \citet{Newberg;et-al_2010} found the Orphan Stream to have $V_{GSR} = 110$ km s$^{-1}$ from BHB stars, approximately $105$\,km s$^{-1}$ on our scale. To isolate potential Orphan Stream members we have nominated a relatively wide kinematic selection region between $75 \leq V_{GSR} \leq 135$ km s$^{-1}$, which yields twenty-five Orphan Stream candidates. The typical uncertainty in our velocities is $\pm{}5.0$\,km s$^{-1}$, although for our fainter targets this quickly increases above 10\,km s$^{-1}$.

@@

We have measured the equivalent width of the gravity-sensitive Mg\,\textsc{I} line at 8807 \AA{} to distinguish dwarfs from giants \citep{Battaglia;Starkenburg_2012}. At a given temperature (or $g - r$) and metallicity, giant stars present narrower Mg\,\textsc{I} absorption lines than their dwarf contaminants. Given our target selection, our sample is likely to contain many more dwarfs than giants. In some cases no Mg\,\textsc{I} 8807 \AA{} line was apparent, so an upper limit was estimated based on the S/N. In these cases the candidate was considered a ``non-dwarf" because we cannot exclusively rule out a metal-poor sub-giant with this criterion alone. For these purposes we are only looking for a simple indication as to whether a star is likely a dwarf or not. Figure \ref{fig:ew-mg} illustrates the trend with EW$_{\lambda8807}$ against SDSS de-reddened\footnote{All magnitudes presented in this letter are de-reddened using the \citet{Schlegel;Finkbeiner;Davis_1998} dust maps.} $g - r$, illustrating the dominant upper dwarf branch we wish to exclude. The adopted dwarf/giant separation line is shown. Admittedly this selection criterion is not particularly strict, but we are allowing multiple inclusive cuts and taking the intersection of these criteria to identify Orphan Stream giants.

@@

We have measured the equivalent widths of the Ca\,\textsc{II} near-infrared triplet lines for stars that meet our kinematic and surface gravity criteria. After correcting for luminosity, the equivalent width of the Ca\,\textsc{II} triplet lines provide a good indication of the overall metallicity of a RGB star \citep{Amandroff;Da_Costa_1991}. We have employed the \citet{Starkenburg;et-al_2010} relationship and corrected for luminosity in $g$ against the horizontal branch magnitude at $g_{HB}$ = 17.1 \citep{Newberg;et-al_2010}. Strictly speaking, the Ca\,\textsc{II}\---[Fe/H] calibration is only valid for stars brighter than the horizontal branch, although the relationship only becomes significantly inappropriate near $g \-- g_{HB} \sim +1$ \citep{Saviane;et-al_2012}. Many of our candidates are fainter than this valid luminosity range, and therefore they should not be excluded solely because of their derived metallicities. Stars fainted than $g_{HB}$ will have slightly lower metallicities than predicted by our Ca\,\textsc{II}\---[Fe/H] relationship, and for these stars we will only use metallicities to assign a relative qualitative likelihood for stream membership.
Given a distance to the Orphan Stream, we can also estimate a star's metallicity through isochrone fitting. We have used a 10\,Gyr \citet{Girardi;et-al_2008} isochrone at 21.4\,kpc \citep{Newberg;et-al_2010} and found metallicities for all eighteen likely stream members from their best-fitting isochrone. Derived metallicities from Ca\,\textsc{II} line strengths and isochrone fitting that are consistent (within $\pm0.3$\,dex) indicates these measurements are reliable, and that these stars are indeed at a distance of $\sim$21.4\,kpc. We find ten highly likely stream members with consistently derived metallicities. They fall within the shaded region illustrated in Figure \ref{fig:feh}.

We have then investigated the status of these ten stars in the PPMXL proper motion catalogue \citep{Roeser;et-al_2010}, finding entries for nine of the candidates. One star (OSS-9 in Table \ref{tab:oss-members}) has a listed proper motion that is different from that of the others at the $6\,\sigma$ level. Consequently, we have reduced the membership likelihood of this star from ``High" to ``Medium".

We have adopted a final overall metallicity for each star based on the quality of our [Fe/H] measurements. Adopted values are tabulated in Table \ref{tab:oss-members}. From our nine highly likely stream members we find an overall stream metallicity of $[\mbox{Fe/H}] = -1.63$ with a dispersion of $\sigma = 0.56$ dex. This abundance spread is larger than typically seen in globular cluster stars and is more representative of the chemical spreads seen in dSph satellites. \citet{Newberg;et-al_2010} found a stream metallicity of $[\mbox{Fe/H}] = -2.1$ from BHB stars, which is not inconsistent with our measurement given the spread in Orphan Stream K-giant metallicities.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/feh.pdf}
	\caption{Metallicities from the Ca\,\textsc{II} triplet lines versus those found from fitting isochrones to the eighteen stars that meet our kinematic and surface gravity criteria. Both abundance determinations imply these stars are RGB members of the Orphan Stream at a distance of $\sim$21.4\,kpc \citep{Newberg;et-al_2010}. Consistency between these methods indicates highly likely stream membership (green region). The minimum isochrone [Fe/H], and a representative uncertainty of 0.2\,dex for abundance measurements is shown.}
	\label{fig:feh}
\end{figure}

Given that our Orphan Stream giants cover a wide evolutionary range along the giant branch (Figure \ref{fig:cmd}), we are in a good position to revise the distance estimate to the stream. Given a 10\,Gyr \citet{Girardi;et-al_2008} isochrone at $[\mbox{Fe/H}] = -1.63$, we find a best-fitting distance to the stream of $22.5 \pm 1.0$\,kpc at $(l, b) = (250\,^\circ,\,50\,^\circ)$. This isochrone is shown in Figure \ref{fig:cmd}, illustrating that three of our nine most probable member stars lie perfectly on this isochrone. Our derived distance is in reasonably good agreement with the $21.4 \pm 1.0$\,kpc deduced by \citet{Newberg;et-al_2010}.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/cmd.pdf}
	\caption{Color magnitude diagram showing our observed candidates (grey). Observations fulfilling kinematic and gravity cuts are colored by their metallicity, and those with upper limits for surface gravity are marked as triangles ($\triangledown$). Highly probable stream members (see text) are circled. Relevant 10\,Gyr \citet{Girardi;et-al_2008} isochrones at $[\mbox{Fe/H}] = -1.5$ (dotted), $-2.0$ (dashed) at 21.4\,kpc \citep{Newberg;et-al_2010} are shown, as well as our best-fitting 10\,Gyr isochrone of $[\mbox{Fe/H}] = -1.63$ at 22.5\,kpc (solid).}
	\label{fig:cmd}
\end{figure}

@@

\bibliographystyle{apj}
\bibliography{bibliography}
